---
title: "COVID-19 socioeconomic impacts"
author: "Mohamed Hachaichi"
date: '2022-03-03'
output: pdf_document
---

### Read the libs 
```{r error=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(readxl)
library(reshape2)
library(modelsummary)


# load package
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```


### Read the data 

```{r error=FALSE}
# read the data 
ds <- read_excel('../Data/Analysis/DS.xlsx')

# drop nan
ds <- ds %>% 
  drop_na()


# add the zscore 
ds <- ds %>%
  mutate(zscore = (ds$Death_rate - mean(ds$Death_rate) / sd(ds$Death_rate))) %>%
  as_data_frame()
  

# Create a dummy variable to identify the group exposed to the treatment

ds$treated = ifelse(ds$zscore >= 5, 1, 0) 

# create the datasets for unemployment and youth unemployment 

## 1 unemployment 

unem <- ds %>%
  select(country, NUTS_ID, Unemployment_2020, Death_rate, Unemployment_2019, 'URBAN-RURAL',Border, METRO, zscore, treated) %>%
  rename('2020' = Unemployment_2020,
         '2019' = Unemployment_2019)


unem <- melt(unem, id.vars = c('country', 'NUTS_ID', 'Death_rate', 'URBAN-RURAL', 'Border', 'METRO', 'zscore', "treated"), variable.name = 'Date')


## Create a dummy variable to indicate the time when the treatment started
unem$time <- ifelse(unem$Date == 2020, 1, 0)

## 2 youth 

youth <- ds %>%
  select(country, NUTS_ID, Youth_2020, Death_rate, Youth_2019, "URBAN-RURAL",Border, METRO, zscore, treated) %>%
  rename('2020' = Youth_2020,
         '2019' = Youth_2019)

youth <- melt(youth, id.vars = c('country', 'NUTS_ID', 'Death_rate', 'URBAN-RURAL', 'Border', 'METRO', 'zscore', "treated"), variable.name = 'Date')
## Create a dummy variable to indicate the time when the treatment started

youth$time = ifelse(youth$Date == 2020, 1, 0)


```


```{r}
ds %>%
  group_by(treated) %>%
  count()

```


```{r}
hist(ds$zscore)
```


```{r}

hist(ds$Unemployment_2020)

```
```{r}
hist(ds$Unemployment_2019)
```

```{r}
hist(ds$Youth_2019)

```

```{r}
hist(ds$Youth_2020)
```



```{r}
unem %>%
  group_by(Date) %>%
  summarise(mean = mean(value, na.rm = TRUE))

```

```{r}
unem %>%
  group_by(Date, treated) %>%
  summarise(mean = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = Date, y = mean, color = treated)) +
  geom_point()

```


```{r}
youth %>%
  group_by(Date) %>%
  summarise(mean = mean(value, na.rm = TRUE))

```


### DiD analysis 

key assumption: zscore(death rate) >= 0 are the regions that were treated.

#### Unemployment 

```{r, error= FALSE, message= FALSE}

total_unem = lm(value ~ treated*time, data = unem)

urban_unem = lm(value ~ treated*time, data = unem %>% filter(unem$`URBAN-RURAL` == 1))

rural_unem = lm(value ~ treated*time, data = unem %>% filter(unem$`URBAN-RURAL` == 3))

intermediate_unem = lm(value ~ treated*time, data = unem %>% filter(unem$`URBAN-RURAL` == 2))

border_unem = lm(value ~ treated*time, data = unem %>% filter(unem$Border == 1))

metro_unem = lm(value ~ treated*time, data = unem %>% filter(unem$METRO == 1))

summary(border_unem)

```

This model indicates a difference-in-differences estimate of 20% lower unemployment growth for those hit by the pandemic (eligible) than those who were not severely hit by the pandemic.

OR: The pandemic causal effect of the pandemic it causes an increase of 20% in unemployment rates (%), 



#### Youth Unemployment 


```{r, error= FALSE, message= FALSE}

total_youth = lm(value ~ treated*time, data = youth)

urban_youth = lm(value ~ treated*time, data = youth %>% filter(youth$`URBAN-RURAL` == 1))

rural_youth = lm(value ~ treated*time, data = youth %>% filter(youth$`URBAN-RURAL` == 3))

intermediate_youth = lm(value ~ treated*time, data = youth %>% filter(youth$`URBAN-RURAL` == 2))

border_youth = lm(value ~ treated*time, data = youth %>% filter(youth$Border == 1))

metro_youth = lm(value ~ treated*time, data = youth %>% filter(youth$METRO == 1))

summary(metro_youth)


```




#### all the models togother 

```{r}

tab_model(total_unem,total_youth)


```


### 2SLS 

```{r}



```

